---
title: "Hams in the USA"
author: "Donna A. Dietz (N2SZ)" 
date: "May 2021"
output: 
  html_document:
    number_sections: TRUE
---

This little data mapping project is in response to the question posed at
[the HRO website](https://www.hamradio.com/newstore.cfm) which is where
to place a new HRO store!

In other words, it's a good excuse to pull out R and do a little data wrangling, 
but in the end it's all just for fun.  I really think the new location should
be in Dayton, Ohio, because that's where all the hams go every year, and there 
isn't one there yet!

But, just for starters, I downloaded the FCC database and pulled out all the
zipcodes for every ham in the database, mapped that to the counties (farily accurate but not perfect), and then created a heatmap of the USA to show where the hams
all live.

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#Joining 3 files for project
#https://cran.r-project.org/web/packages/usmap/vignettes/advanced-mapping.html

library(tidyverse)

library(usmap)
EN <- read_csv("EN.csv")  #FCC
#https://www.fcc.gov/uls/transactions/daily-weekly#fcc-uls-transaction-files-weekly  
#(l_amat/EN.dat)


ZipCounty <- read_csv("ZIP-COUNTY-FIPS_2017-06.csv")  #Kaggle
#https://www.kaggle.com/danofer/zipcodes-county-fips-crosswalk

EN <- EN %>% mutate(zip=substring(`'zip'`,2,6))

EN <- EN %>% mutate(zip2=as.numeric(zip))

mydata <- left_join(EN,ZipCounty, by=c("zip2"="ZIP"))

hamCount <- mydata %>% group_by(STCOUNTYFP) %>% count()

county_fips <- read_csv("county_fips_master.csv")
#https://github.com/kjhealy/fips-codes/blob/master/county_fips_master.csv




#library(usmap)
hamCount$fips <-hamCount$STCOUNTYFP
hamCount$countyFIPS <- as.numeric(hamCount$STCOUNTYFP)

hamCount2 <-left_join(county_fips, hamCount, by=c("fips"="countyFIPS"))

hamCount3 <- hamCount2 %>% mutate(n=replace_na(n,0))
```

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
plot_usmap(regions=c("counties"), data=hamCount3, values="n")+theme(legend.position="right")+ 
  scale_fill_continuous(name="Hams per County", label=scales::comma)+
  labs(title="Amateur Radio Operators in the United States",
       subtitle="(California is REALLY messing up this graphic!)")
```

So, Southern California is loaded with hams.  Let's just take the log of this
data so we can get another viewpoint, shall we? (Just for fun!)

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
hamCount4 <- hamCount3 %>% mutate(logdata = log(`n`+1))
plot_usmap(regions=c("counties"), data=hamCount4, values="logdata")+theme(legend.position="right")+ 
  scale_fill_continuous(name="ln(Hams per County)", label=scales::comma)+
  labs(title="log of (Amateur Radio Operators in the United States) ",
       subtitle="(Give the rest of us a chance, California!)")

```


But what we really want to know is where to put a new store, right? So let's group some of the states together and plot the locations of the current stores.  I'm starting with my home region.  I might just suggest that the new store should be somewhere near DC, 
and this would probably not be bad at all!  Lots of "smart people" travel to DC on business and might be happy to take a side trip.  I know there is already a store in Virginia, but it seriously would take 2 to 3 hours to get to it from Fort Meade depending on the traffic.  That 95 corridor can get really horrific!

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
# NORTH EAST SECTION
#home <- tibble(lon=-77, lat=39)
#homeTrans<-usmap_transform(home)
#stores in NewCastle, Woodbridge, Salem
hroNE <- tibble(lon=c(-75.6,-77.3,-71.2), lat=c(39.7, 38.6, 42.8))
hroNE_Trans <- usmap_transform(hroNE)

plot_usmap(regions=c("counties"), 
           include=c("NJ","PA","DE", "MD","DC","VA","CT","NY","MA",
                     "WV","NH","VT","ME","OH","NC"), 
           data=hamCount, values="n")+ theme(legend.position="right")+ 
  scale_fill_continuous(name="Hams per County", label=scales::comma)+
  geom_point(data = hroNE_Trans, aes(x = lon.1, y = lat.1),
             color = "red", alpha = 0.8)+
  labs(title="Amateur Radio Operators in the Eastern United States",
       subtitle="HRO Stores: Salem-NH, New Castle-DE, Woodbridge-VA")

```



It was so much fun making that map, I thought I should group other states together to show off their stores!  Here are some in the southern US.


```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#TEXAS ETC SECTION  #STORES in Denver, Phoenix, Plano
hroCO <- tibble(lon=c(-104.9,-112.2,-96.7), lat=c(39.7,33.6,33.0))
hroCO_Trans <- usmap_transform(hroCO)
plot_usmap(regions=c("counties"), 
           include=c("NM","CO","AZ","TX"), 
           na=0, data=hamCount3, values="n")+ theme(legend.position="right")+ 
  scale_fill_continuous(name="Hams per County", label=scales::comma)+
  geom_point(data = hroCO_Trans, aes(x = lon.1, y = lat.1),
             color = "red", alpha = 0.8)+
  labs(title="Amateur Radio Operators in Texas, Arizona, NM, CO",
       subtitle="HRO Stores: Denver-CO, Phoenix-AZ, Plano-TX")


```



Of course, California!

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#Cali SECTION  #STORES in Anaheim, Oakland, SanDiego
hroCA <- tibble(lon=c(-117.9,-122.2,-117.1), lat=c(33.8,37.8,32.8))
hroCA_Trans <- usmap_transform(hroCA)
plot_usmap(regions=c("counties"), 
           include=c("CA"), 
           na=0, data=hamCount3, values="n")+ theme(legend.position="right")+ 
  scale_fill_continuous(name="Hams per County", label=scales::comma)+
  geom_point(data = hroCA_Trans, aes(x = lon.1, y = lat.1),
             color = "red", alpha = 0.8)+
  labs(title="Amateur Radio Operators in California",
       subtitle="HRO Stores: Oakland (San Fran), Anaheim (LA), San Diego")



```



Oregon has quite a few hams!

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#Portland Oregon store and state
hroOR <- tibble(lon=c(-122.8), lat=c(45.4))
hroOR_Trans <- usmap_transform(hroOR)
plot_usmap(regions=c("counties"), 
           include=c("OR"), 
            data=hamCount3, values="n")+ theme(legend.position="right")+ 
  scale_fill_continuous(name="Hams per County", label=scales::comma)+
  geom_point(data = hroOR_Trans, aes(x = lon.1, y = lat.1),
             color = "red", alpha = 0.8)+
  labs(title="Amateur Radio Operators in Oregon", subtitle="HRO Store: Portland")



```


And so does Milwaukee!

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#Milwaukee WI store
hroWI <- tibble(lon=c(-88.0), lat=c(43.1))
hroWI_Trans <- usmap_transform(hroWI)
plot_usmap(regions=c("counties"), 
           include=c("WI"), 
           data=hamCount3, values="n")+ theme(legend.position="right")+ 
  scale_fill_continuous(name="Hams per County", label=scales::comma)+
  geom_point(data = hroWI_Trans, aes(x = lon.1, y = lat.1),
             color = "red", alpha = 0.8)+
  labs(title="Amateur Radio Operators in Wisconsin",subtitle="HRO Store: Milwaukee")

```


Also Atlanta.

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#Atlanta GA store
hroGA <- tibble(lon=c(-84.3), lat=c(33.9))
hroGA_Trans <- usmap_transform(hroGA)
plot_usmap(regions=c("counties"), 
           include=c("GA"), 
           data=hamCount3, values="n")+ theme(legend.position="right")+ 
  scale_fill_continuous(name="Hams per County", label=scales::comma)+
  geom_point(data = hroGA_Trans, aes(x = lon.1, y = lat.1),
             color = "red", alpha = 0.8)+
  labs(title="Amateur Radio Operators in Georgia",subtitle="HRO Store: Atlanta")

#knitr::spin("mappingTheUS_Hams.R",FALSE,format="Rmd")
```


So I hope you enjoyed looking at my graphics.  I had fun making them.  But in reality, I think there should be a new store either in Dayton, Ohio, or at least put one near Pittsburgh in a place where it would be convenient for (many of us) to stop off on our way to or from Hamvention.


